---
title: "Coding Challenge 7"
author: "DJT"
date: "2026-03-26"
output:
  md_document:
    variant: gfm
  html_document:
    theme: sandstone
    toc: true  
    toc_depth: 3 
    toc_float: true
    code_folding: hide
  word_document:
  pdf_document:
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(knitr)
library(drc)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
```

# Question 1
## Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
### See previous code section for loaded libraries.

```{r}
PlantEmergence <- read.csv("PlantEmergence.csv")
head(PlantEmergence)

#saving columns into factors
PlantEmergence$Treatment <- as.factor(PlantEmergence$Treatment) 
PlantEmergence$DaysAfterPlanting <- as.factor(PlantEmergence$DaysAfterPlanting)
PlantEmergence$Rep <- as.factor(PlantEmergence$Rep)

```

# Question 2
## Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results.

```{r}
lm.interaction.emergence <- lm(Emergence ~ Treatment * DaysAfterPlanting, data = PlantEmergence) #linear regression with interaction via *

summary(lm.interaction.emergence) #Model summary

anova(lm.interaction.emergence) #ANOVA table

```
# Question 3
## Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.
### Based on the results from question 2 you do not need to fit the interaction term, it had a p value of 1. Treatment 2 has, on average, 134.53 fewer emerged plants than Treatment 1, holding DaysAfterPlanting constant with a p value less than .05.

```{r}
lm.simple.emergence <- lm(Emergence ~ Treatment+DaysAfterPlanting, data = PlantEmergence) #simple linear regression with +

summary(lm.simple.emergence)

anova(lm.simple.emergence)
```
# Question 4
## Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r}
lsmeans <- emmeans(lm.interaction.emergence, ~Treatment) # calculating the least sqaure means and saving it as an object
lsmeans

results_lsmeans <- cld(lsmeans, alpha = 0.05, details = TRUE)
results_lsmeans
```
# Question 5
## The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters.
### The letter indicates the statistical significance between the treatments. Shared letters represent non significance whereas unique letters do as indicated by the p value less than .05. Treatment 2 is significantly different than all other treatments and treatment 7 is significantly different from 3, 5, and 6.

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}
plot_cldbars_onefactor(lm.simple.emergence, "Treatment")
```

# Link to dzt0062 git
[CC7 - gfm]([https://github.com/dzt0062/PLPA_assignments/blob/main/Coding_Challenges/Coding_Challenge_7_files/CC7.Rm)

