---
title: "Coding Challenge 5"
author: "DJT"
date: "2026-03-05"
output:
  html_document:
    theme: sandstone
    toc: true          
    toc_depth: 3       
    toc_float: true
    code_folding: hide
  md_document:
    variant: gfm
  word_document:
  pdf_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(knitr)
```

# Question 1
## Download two .csv files from Canvas called DiversityData.csv and Metadata.csv, and read them into R using relative file paths.

```{r}
setwd("~/PLPA_assignments")
data.meta<-read.csv("PLPA_assignments/Data/Metadata.csv", na.strings = "na")
data.div<-read.csv("PLPA_assignments/Data/Diversitydata.csv", na.strings = "na")
```

# Question 2
## Join the two dataframes together by the common column ‘Code’. Name the resulting dataframe alpha.

```{r}
alpha<-left_join(data.div, data.meta, by = "Code") #joining the dataframes
head(alpha)
```

# Question 3
## Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column
### a.	Using mutate, create a new column to calculate Pielou’s evenness index.
### b.	Name the resulting dataframe alpha_even.

```{r}
alpha_even <- alpha %>%
  mutate(alpha, pielou = shannon/log(richness)) #new dataframe called alpha_even with a new column calculating Pielou's evenness
head(alpha_even)
```

# Question 4
## Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.
### a.	Start with the alpha_even dataframe 
### b.	Group the data: group the data by Crop and Time_Point.
### c.	Summarize the data: Calculate the mean, count, standard deviation, and standard error for the even variable within each group.
### d.	Name the resulting dataframe alpha_average

```{r}
alpha_average<-alpha_even %>%
  group_by(Crop, Time_Point)%>% # grouped by Crop and Time Pint
  summarise(Mean.pielou = mean(pielou), #stats via summarize 
            n = n(), 
            sd.dev = sd(pielou)) %>%
  mutate(std.err = sd.dev/sqrt(n)) #adding a column for standard dev.
head(alpha_average)
alpha_average
```

# Question 5
## Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
### a.	Start with the alpha_average dataframe
### b.	Select relevant columns: select the columns Time_Point, Crop, and mean.even.
### c.	Reshape the data: Use the pivot_wider function to transform the data from long to wide format, creating new columns for each Crop with values from mean.even.
### d.	Calculate differences: Create new columns named diff.cotton.even and diff.soybean.even by calculating the difference between Soil and Cotton, and Soil and Soybean, respectively.
### e.	Name the resulting dataframe alpha_average2

```{r}
alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, Mean.pielou)%>% # Select relevant columns
  pivot_wider(names_from = Crop, values_from = Mean.pielou)%>% # Reshape the data creating new columns for each Crop with values from mean.even
  mutate(diff.cotton.even = Soil - Cotton, diff.soybean.even = Soil - Soybean) #Calculated difference between two new columns named diff.cotton.even and diff.soybean.even
head(alpha_average2)
```

#Question 6
## Connecting it to plots
### a.	Start with the alpha_average2 dataframe
### b.	Select relevant columns: select the columns Time_Point, diff.cotton.even, and diff.soybean.even.
### c.	Reshape the data: Use the pivot_longer function to transform the data from wide to long format, creating a new column named diff that contains the values from diff.cotton.even and diff.soybean.even. ***This might be challenging, so I’ll give you a break. The code is: pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff")***
### d.	Create the plot: Use ggplot and geom_line() with ‘Time_Point’ on the x-axis, the column ‘values’ on the y-axis, and different colors for each ‘diff’ category. The column named ‘values’ come from the pivot_longer. The resulting plot should look like the one to the right.

```{r}
alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even)%>% #Select relevant columns
   pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff", values_to = "values")%>% # Reshape the data to long form with new column named diff that contains the values from diff.cotton.even and diff.soybean.even.
    ggplot(aes(x = Time_Point, y = values, color = diff, group = diff))+ #ggplot with aes layering and classic theme
    geom_line()+
    theme_classic()+
    xlab("Time (hrs)") +
    ylab("Difference from soil in Pielou's eveness")
```


# Question 7
## Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 5. Provide me a link to your github written as a clickable link in your .pdf or .docx


